---
title: "CHU_H Final Poject Step One Data Acquisition"
output: html_notebook
---

This is the library we need for this phase of the project.

```{r}
library(jsonlite)
```

To obtain our data, we will need to perform a call to the official osu api.

```{r}
# This is the base url for the api, all api calls will have this in the beginning
base.api.url <- "https://osu.ppy.sh/api/"

# This is the parameter that gets beatmap data from the api
get.beatmap.request <- "get_beatmaps?"

# This is my personal api key, please use your own
personal.api.key <- "k=725913db82908989f75b458d76f36ded87176eec"

# The parameter that tells the api to return all beatmaps made from since April 1st, 2018
since.parameter <- "&since=2018-04-01"

# Since we are only focusing on osu!standard mode the parameter would be 0
mode.parameter <- "&m=0"

# The final url we are using to call the api
final.request.url <- paste(base.api.url,get.beatmap.request,personal.api.key,since.parameter,mode.parameter, sep="")
```

Now we will make a function that would make the call and returns the data obtained in a dataframe

```{r}
# Returns a dataframe with the info from the 500 latest beatmaps
beatmap.data.since.april <- fromJSON(txt = final.request.url)
print(beatmap.data.since.april)
```

It is important to note that this data is collected on 4.15.18 8:00PM because beatmaps are constantly being uploaded and these data will only be true up until this point in time. To preserve our data we can store them into a CSV file.

```{r}
write.table(beatmap.data.since.april, "mostRecent500BeatmapStatisticApril.csv",row.names = FALSE, sep = ",")
```

